Update and evaluation of the ozone dry deposition in the OsloCTM3 v1.0

### 2018-12-05 DryDep.tex
Updated dry deposition velocity calculation and corresponding number in the text. Need to go through the whole of it again.

Average dry deposition velocities have been reconstructed from the ratio between surface concentration and total dry deposition
\begin{equation}
  v^\chem{O_3}_\text{DD} = \frac{\sum \chem{O_3^\text{DD}}(z_0)}{\chem{O_3}(z_0)}\cdot c_\text{month},
\end{equation}
with $c_\text{month} = \frac{\text{layer thickness}}{\text{seconds in month}}$. This way of retrospectively computing dry deposition velocities has certain drawbacks of which one is the freedom of choice of the \emph{layer thickness}. We chose the average thickness of the lowermost model level ($16\cdot10^{2}\,\unit{cm}$) to match the prescribed dry deposition velocities above ice and snow ($v_\text{ice/snow}^\chem{O_3} = 0.05\,\unit{cm\,s^{-1}}$).
%
\subsubsection{Zonal distribution}
\label{subsubsec:zonal}
%
The annual zonal average with respect to surface ozone concentration (Fig.~\ref{fig:mmm_drydep}a) displays, in consistency with Fig.~\ref{fig:diff_maps}a), a general, global increase of ozone by up to $10\,\unit{ppb}$. This increase is largest in the zonal band $(25-50)\,\unit{^\circ N}$ which contains the major deserts. In the deep tropics ($5\,\unit{^\circ S}-5\,\unit{^\circ N}$), the increase is smallest ($\Delta[\chem{O_3}]\leq 5\,\unit{ppb}$). We find that the EMEP scheme further intensifies the strong asymmetry between northern and southern hemisphere as a consequence of the distribution of the continental land masses and vegetation thereon. Among the sensitivity studies, there is only a low absolute variance. Most remarkable but expected due to the much smaller prescribed dry deposition velocity over ice and snow, \emph{EMEP\_ppgssh\_ice} displays almost a doubling of surface ozone in the high Arctics compared to \emph{Wesely\_type} but affects ozone concentrations down to latitudes at about $50\,\unit{^\circ}$ in both hemispheres. Neglecting any dependence on solar radiation in the stomatal conductance formulation (\emph{EMEP\_offLight}) -- or in other words allowing photosynthesis 24/7 -- generally decreases the ozone concentration by about $2\,\unit{ppb}$, while choosing soil water at deeper levels than the surface layer (\emph{EMEP\_SWVL4}) increases $[\chem{O_3}](z_0)$ slightly. Rather surprisingly, switching off the phenology completely (\emph{EMEP\_offPhen}), amounts only to small difference (at most $3.5\,\unit{\%}$ increase in the deep tropics). Switching from the SMOKE-BEIS parameterization to the pre-processed greening season (\emph{EMEP\_ppgs} compared to \emph{EMEP\_full}), surface ozone increases on average by $6\,\unit{\%}$ between $54-70\,\unit{^\circ N}$. The scenario of differing emissions (2005 in comparison to 2014 or more specifically \emph{EMEP\_ppgs\_2005} compared to \emph{EMEP\_ppgs}), yields higher ozone concentrations in the northern hemisphere in 2005 in accordance to a reduction in sulfur and \chem{NO_x} emissions in south east Asia in later years. An opposite tendency is seen for latitudes south of $30\,\unit{^\circ N}$, probably owing to the industrial development of countries in these regions.

The $v^\chem{O_3}_\text{DD}$ are shown in Fig.~\ref{fig:mmm_drydep}b). The dry deposition velocities in the EMEP scheme are much smaller than in the Wesely scheme for latitudes between $50\,\unit{^\circ S}$ and $70\,\unit{^\circ N}$. In the Arctics, except for \emph{EMEP\_ppgssh\_ice}, all realizations of the EMEP scheme are constantly above the multi-model-mean. This indicates, that with respect to other models, the updated dry deposition velocity above ice and snow should be considered as new standard for the OsloCTM3. The dry deposition velocities are of course independent of the emission scenario, but display a strong dependence on $f_\text{light}$ and $f_\text{phen}$. The closer to $1$ these factors are the higher the dry deposition velocities become (compare \emph{EMEP\_offLight}, \emph{EMEP\_offPhen}, and \emph{EMEP\_ppgssh}). That highlights the importance of a proper PAR from meteorological input in CTMs or radiative transfer in PAR in global climate models (GCMs) for canopy conductance computation. Albeit, the total numbers are smaller, the shape of the normalized zonal average dry deposition velocities of the EMEP scheme and the multi-model-mean are similar (Supplement~S.6.1). The biggest exceptions are the zonal bands $(50-70)\,\unit{^\circ S}$, which is almost entirely covered by ocean, $(12-30)\,\unit{^\circ S}$, which coincides with the location of Australia and its desert regions, as well as its counterpart in the northern hemisphere $(12-30)\,\unit{^\circ N}$. However, it is not clear whether diverging dry deposition velocities over ocean or deserts have the larger impact.

@@

\begin{figure}[t]
  \includegraphics[width=8.3cm]{pictures/final-vmr_surfaceozone}\\
  \includegraphics[width=8.3cm]{pictures/final-total_ozone_drydepvelo}\\
  \includegraphics[width=8.3cm]{pictures/final-total_ozone_drydep}
  \caption{Comparison of the manifold of OsloCTM3 integrations with respect to (a) Surface ozone concentrations, (b) Annual average ozone dry deposition velocity, (c) Total annual ozone dry deposition. The different colors indicate sets of simulation with similar baselines. The multi-model mean from the evaluation of HTAP models by \citet{ACP:Hardacre2015} is shown as a reference (where available).}
  \label{fig:mmm_drydep}
\end{figure}
%
In Fig.~\ref{fig:mmm_drydep_season}a), the average zonal ozone dry deposition is shown separated by month. Where available, we have added the multi-model-mean given by \citet{ACP:Hardacre2015} as reference. There are two major phases apparent: NH and SH greening season. Spring and summer in the NH is reflected in a pronounced peak of $v_\text{DD}^\chem{O_3}$ in the northern mid latitudes, while it is absent in winter (SH summer). Spring and summer in the SH are marked by a southward shift of the tropical peak dry deposition velocity. In the Wesely scheme, NH mid latitude peak velocities appear in June compared to July in the EMEP scheme, indicating that the seasonal cycles differ. The corresponding total monthly ozone dry deposition is shown in Fig.~\ref{fig:mmm_drydep_season}b). In general, the seasonal patterns are quite similar in the Wesely scheme and the EMEP scheme, displaying a strong symmetry around $10\,\unit{^\circ N}$ in January/February and November/December, respectively. What differs most is the molding and intensity of the NH peak dry deposition. Both schemes reach the maximum in June/July but the peak is much more differentiated in March already in the Wesely scheme. Similarly, the SH tropical peak dry deposition is reached in August/September but sustained longer, into October, in the Wesely scheme. Since we have not conducted any simulation with a meteorological year other than 2005, we cannot elaborate on whether this is a special feature of our chosen year or not. In comparison, the available two month of the multi-model-mean display a different pattern. Especially the NH/SH asymmetry is much more pronounced in February, than it is the EMEP scheme.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Average seasonal cycles}
\label{subsubsec:seasons}
%
To further disentangle the contributions of different regions to the global ozone budget, we will briefly look at different projections of seasonal cycles.

@@

In Fig.~\ref{fig:mmm_drydep_season_pft}, the seasonal cycles of dry deposition velocities are shown for the nine surface categories. The patterns and absolute numbers differ substantially between the Wesely scheme and the EMEP scheme and the multi-model-mean. The divergence of the average dry deposition velocities between the Wesely scheme and the EMEP scheme in desert regions ($\Delta \overline{v}_\text{desert}^\chem{O_3} = 0.14\,\unit{cm\,s^{-1}}$) as well as grassland ($\Delta \overline{v}_\text{grassland}^\chem{O_3} = 0.4\,\unit{cm\,s^{-1}}$) is quite remarkable as well as the difference of both to the multi-model-mean in tropical forest regions ($\Delta \overline{v}_\text{tropical forest}^\chem{O_3} = 0.38\,\unit{cm\,s^{-1}}$). In contrast to the Wesely scheme and the EMEP scheme, the multi-model-mean indicates a rather pronounced seasonal cycle in desert regions ($0.10\,\unit{cm\,s^{-1}} \leq v^\chem{O_3}_\text{desert}\leq 0.15\,\unit{cm\,s^{-1}}$). The estimated dry deposition velocities over desert regions are consistent with the previously mentioned baseline values of ozone dry deposition, which means that in the EMEP scheme they are about one order of magnitude lower than in the Wesely scheme. From measurements of ozone fluxes in the Sahara desert, \citet{AE:Gusten1995} deduced $v^\chem{O_3}_\text{desert, day} = 0.1\,\unit{cm\,s^{-1}}$, $v^\chem{O_3}_\text{desert, night} = 0.04\,\unit{cm\,s^{-1}}$, and $\overline{v}^\chem{O_3}_\text{desert} = 0.065\,\unit{cm\,s^{-1}}$. This implies, that ozone dry deposition over desert regions is highly overestimated in the Wesely scheme as well as in the multi-model-mean, while it may be underestimated in the EMEP scheme. Similarly, the dry deposition velocities over water differ. From measurements over the ocean a mean value of $\overline{v}^\chem{O_3}_\text{water} = 0.019\,\unit{cm\,s^{-1}}$ has been deduced \citep{JGR:Helmig2012}. In a model study of different mechanisms of dry deposition to ocean waters by means of prescribed $\overline{v}_\text{water}^\chem{O_3}$ and one- and two-layer gas exchange modeling, \citet{ACP:Luhar2017} found $\overline{v}_\text{water}^\chem{O_3}$ ranging between $0.018\,\unit{cm\,s^{-1}}$ (two-layer scheme) and $0.039\,\unit{cm\,s^{-1}}$ (prescribed). With $\overline{v}^\chem{O_3}_\text{water} = (0.038\pm0.002)\,\unit{cm\,s^{-1}}$, the EMEP scheme (Section~\ref{subsubsec:Rb}) yields probably a too strong dry deposition to ocean which implicates that ozone concentrations might even become larger and dry deposition even lower in the model if a two-layer scheme would be implemented. With respect to forests, we might be able to improve the model performance further by allowing more PFTs, instead of using a broad average. 

@@

Finally, we have a look at the different global as well as hemispheric sinks for ozone. Despite its vastness, the ocean amounts only to about $23\,\unit{\%}$ of the global ozone sink in the OsloCTM3 with operative EMEP scheme, while ice and snow account for far less than $1\,\unit{\%}$. The remainder is deposited to various other land surfaces of which deserts might be the most neglected in process modeling. The total annual dry deposition in the EMEP scheme is about $\frac{1}{3}$ below the multi-model-mean result by \citet{ACP:Hardacre2015}. But also the results of ~\citet{ACP:Luhar2017} yield a $(19-27)\,\unit{\%}$ lower ozone dry deposition than the models participating in the model intercomparison, with deposition to ocean ranging between $(12-21)\,\unit{\%}$.
%
\begin{table*}[t]
\caption{Total ozone dry deposition for the respective year in \unit{Tg\,a^{-1}} separated into ocean, ice and, land contributions.}
\begin{tabular}{lccccccccccccr}
\tophline
\multirow{3}{*}{Simulation} & \multicolumn{3}{c}{Ocean} & \multicolumn{3}{c}{Ice} & \multicolumn{3}{c}{Land} & \multicolumn{3}{c}{Total} & $\Delta^\dagger$\\
& NH & SH & Global & NH & SH & Global & NH & SH & Global & NH & SH & Global\\
& \multicolumn{3}{c}{(\unit{Tg\,a^{-1}})} & \multicolumn{3}{c}{(\unit{Tg\,a^{-1}})} & \multicolumn{3}{c}{(\unit{Tg\,a^{-1}})} & \multicolumn{3}{c}{(\unit{Tg\,a^{-1}})} & (\unit{\%})\\
\middlehline
Wesely\_type & 88.7 & 110.3 & 199.0 & 0.7 & 5.0 & 5.7 & 346.1 & 153.5 & 499.6 & 612.4 & 347.9 & 960.2 & 46.8\\
EMEP\_full & 67.1 & 84.7 & 151.8 & 0.6 & 4.4 & 5.0 & 236.0 & 112.3 & 348.2 & 408.6 & 245.6 & 654.2 & 0.0\\
EMEP\_offLight & 66.6 & 84.1 & 150.7 & 0.6 & 4.4 & 5.0 & 267.8 & 129.8 & 397.6 & 451.3 & 267.2 & 718.6 & 9.8\\
EMEP\_offPhen & 67.1 & 84.6 & 151.7 & 0.6 & 4.4 & 5.0 & 239.9 & 116.0 & 355.9 & 413.9 & 249.9 & 663.7 & 1.5\\
EMEP\_SWVL4 & 68.1 & 86.1 & 154.2 & 0.6 & 4.5 & 5.1 & 241.8 & 115.0 & 356.8 & 417.2 & 250.7 & 667.9 & 2.1\\
EMEP\_ppgs & 67.1 & 84.6 & 151.7 & 0.6 & 4.4 & 5.0 & 238.7 & 112.6 & 351.3 & 411.4 & 246.0 & 657.4 & 0.5\\
EMEP\_ppgssh & 67.0 & 84.5 & 151.6 & 0.6 & 4.4 & 5.0 & 243.9 & 119.4 & 363.3 & 419.4 & 253.6 & 673.0 & 2.9\\
EMEP\_ppgssh\_ice & 66.5 & 84.8 & 151.3 & 0.2 & 1.3 & 1.4 & 240.3 & 120.4 & 360.7 & 413.0 & 251.1 & 664.1 & 1.5\\
EMEP\_ppgs\_2005 & 66.8 & 83.5 & 150.4 & 0.6 & 4.4 & 5.0 & 237.8 & 110.0 & 347.8 & 409.2 & 240.7 & 649.9 & -0.7\\
\bottomhline
\end{tabular}
\belowtable{$^\dagger$: Difference of global annual total relative to \emph{EMEP\_full}.} % Table Footnotes
\end{table*}
%

@@

In Fig.~\ref{fig:mmm_drydep_stations}a), seasonal cycles of average ozone dry deposition fluxes for six observation sites are shown. We have computed a model average for all sensitivity studies at the closest grid point and show the $1\,\sigma$ error band. The shaded area around the multi-model-mean indicates the broad range of model results but is not an actual error band since such is not given in \citet{ACP:Hardacre2015}. At about $4$ of $6$ sites, the EMEP scheme performs better than the Wesely scheme and better than or the same as the multi-model-mean. We use a $\chi^2$-test
\begin{equation}
  \chi^2 = \sum_{i=1}^{12}\frac{\left(\overline{\chem{O_3}}_{\text{DD,\,}i}^\text{sim}-\overline{\chem{O_3}}_{\text{DD,\,}i}^\text{obs}\right)^2}{\sigma_i^2},
\end{equation}
with an estimated standard deviation of observation $\sigma_i=1\,\unit{mmol\,m^{-2}\,s^{-1}}$ and divide it by the number degrees of freedom (NDF) to assess this subjective analysis in a more objective way. The closer to $1$ this test scores, the better does the simulation represent the observation. A score between $0$ and $1$ indicates that the estimated $\sigma$ is too small. The results of the $\chi^2$-test are shown together with the divergences in Fig.~\ref{fig:mmm_drydep_stations}. The $\chi^2$-test reveals also that in $4$ of $6$ cases the EMEP scheme improves the performance of the OsloCTM3 with respect to ozone dry deposition fluxes, although a satisfying result is only achieved for two sites (Castel Porziano, Blodgett Forest). In the case of Castel Porziano, Auchencoth Moss, and Citrus Orchard, there may be a hint to an underlying model bias of a too small dry deposition of about $2.8\,\unit{mmol\,m^{-2}\,s^{-1}}$. With only one full year of simulation, the model uncertainty regarding the seasonal cycle at observational sites cannot be quantitatively assessed. Further more, the observational averages uses at most nine years worth of data which questions the statistical reliability of these data. Among other aspects that might affect the model performance the horizontal as well vertical resolution play an important role. In the rather coarse resolution, good matches between observation and model are only to be expected if the station is at a location which is representative for an area similar to the model resolution.
%
\begin{figure}[t]
  \includegraphics[width=8.3cm]{pictures/final-monthly_ozone_drydep_stations_True}\\
  \includegraphics[width=8.3cm]{pictures/final-monthly_drydep_divergence_True}
  \caption{Ozone dry deposition fluxes at different observation sites. (a) Comparison between Wesely scheme and average result from all sensitivity studies with observational averages taken from \citet{ACP:Hardacre2015}. The model uncertainty of the OsloCTM3 is given as $1\,\sigma$ error band. The shaded area around the multi-model-mean indicates the broad range of different model results but is not an actual error band. (b) Model divergence from observation and $\chi^2$-test results.}
  \label{fig:mmm_drydep_stations}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\conclusions[Discussion and conclusions]  %% \conclusions[modified heading if necessary]
\label{sec:conc}
We have presented an update of the dry deposition scheme in the OsloCTM3 from prescribed dry deposition velocities \citep{AE:Wesely1989,JGR:Hough1991} to a more process oriented parameterization taking the state of the atmosphere and vegetation into account \citep{ACP:Simpson2012}. The dry deposition to vegetation is now sensitive to parameters such as photosynthetic active radiation, water vapor pressure deficit, soil water content, and plant phenology. The update decreases the annual amount of ozone dry deposition up to $100\,\unit{\%}$ especially over large desert regions and the oceans. Compared to results from a multi-model evaluation \citep{ACP:Hardacre2015}, the total annual ozone dry deposition in the OsloCTM3 is about $33\,\unit{\%}$ below the average. However, there seems to be a tendency, that the newer TF~HTAP-models show a lower total annual dry deposition of ozone than older models, indicating that newer developments in the models lead to decreasing ozone dry deposition and increasing tropospheric ozone burden \citep[e.g.,][]{ACP:Luhar2017,AE:Hu2017}.

>>>>

For each gridbox, average dry deposition velocities have been reconstructed from the ratio between the monthly averaged  surface ozone concentration and total ozone dry deposition 
\begin{equation}
  v^\chem{O_3}_\text{DD} = \frac{\sum \chem{O_3^\text{DD}}(z_0)}{\chem{O_3}(z_0)}\cdot c_\text{month}.
\end{equation}
Herein $c_\text{month} = \frac{\Delta h_\text{month}}{s_\text{month}}$ with the average height of the lowermost model level in each gridbox $\Delta h_\text{month}$, which is changing with temperature and surface pressure, and the respective number of seconds in a month $s_\text{month}$.
%
\subsubsection{Zonal distribution}
\label{subsubsec:zonal}
%
The annual zonal average with respect to surface ozone concentration (Fig.~\ref{fig:mmm_drydep}a) displays, in consistency with Fig.~\ref{fig:diff_maps}a), a general, global increase of ozone by up to $10\,\unit{ppb}$. This increase is largest in the zonal band $(25-50)\,\unit{^\circ N}$ which contains the major deserts. In the deep tropics ($5\,\unit{^\circ S}-5\,\unit{^\circ N}$), the increase is smallest ($\Delta[\chem{O_3}]\leq 5\,\unit{ppb}$). We find that the EMEP scheme further intensifies the strong asymmetry between northern and southern hemisphere as a consequence of the distribution of the continental land masses and vegetation thereon. Among the sensitivity studies, there is only a low absolute variance. Most remarkable but expected due to the much smaller prescribed dry deposition velocity over ice and snow, \emph{EMEP\_ppgssh\_ice} displays almost a doubling of surface ozone in the high Arctics compared to \emph{Wesely\_type} but affects ozone concentrations down to latitudes at about $50\,\unit{^\circ}$ in both hemispheres. Neglecting any dependence on solar radiation in the stomatal conductance formulation (\emph{EMEP\_offLight}) -- or in other words allowing photosynthesis 24/7 -- generally decreases the ozone concentration by about $2\,\unit{ppb}$, while choosing soil water at deeper levels than the surface layer (\emph{EMEP\_SWVL4}) increases $[\chem{O_3}](z_0)$ slightly. Rather surprisingly, switching off the phenology completely (\emph{EMEP\_offPhen}), amounts only to small difference (at most $3.5\,\unit{\%}$ increase in the deep tropics). Switching from the SMOKE-BEIS parameterization to the pre-processed greening season (\emph{EMEP\_ppgs} compared to \emph{EMEP\_full}), surface ozone increases on average by $6\,\unit{\%}$ between $(54-70)\,\unit{^\circ N}$. The scenario of differing emissions (2005 in comparison to 2014 or more specifically \emph{EMEP\_ppgs\_2005} compared to \emph{EMEP\_ppgs}), yields higher ozone concentrations in the northern hemisphere in 2005 in accordance to a reduction in sulfur and \chem{NO_x} emissions in south east Asia in later years. An opposite tendency is seen for latitudes south of $30\,\unit{^\circ N}$, probably owing to the industrial development of countries in these regions.

The $v^\chem{O_3}_\text{DD}$ are shown in Fig.~\ref{fig:mmm_drydep}b). The dry deposition velocities in the EMEP scheme are on average $43\,\unit{\%}$ smaller than in the Wesely scheme. In the Arctics, except for \emph{EMEP\_ppgssh\_ice}, all realizations of the EMEP scheme are constantly above the multi-model-mean. This indicates, that with respect to other models, the updated dry deposition velocity above ice and snow should be considered as new standard for the OsloCTM3. The dry deposition velocities are of course independent of the emission scenario, but display a strong dependence on $f_\text{light}$ and $f_\text{phen}$. The closer to $1$ these factors are the higher the dry deposition velocities become (compare \emph{EMEP\_offLight}, \emph{EMEP\_offPhen}, and \emph{EMEP\_ppgssh}). That highlights the importance of a proper PAR from meteorological input in CTMs or radiative transfer in PAR in global climate models (GCMs) and a proper choice of phenology for canopy conductance computation. Albeit, the total numbers are smaller, the shape of the normalized zonal average dry deposition velocities of the EMEP scheme and the multi-model-mean are similar (Supplement~S.6.1). The biggest exceptions are the zonal bands $(50-70)\,\unit{^\circ S}$, which is almost entirely covered by ocean, $(12-30)\,\unit{^\circ S}$, which coincides with the location of Australia and its desert regions, as well as its counterpart in the northern hemisphere $(12-30)\,\unit{^\circ N}$. However, it is not clear whether diverging dry deposition velocities over ocean or deserts have the larger impact.

@@

\begin{figure}[t]
  \includegraphics[width=8.3cm]{pictures/final-vmr_surfaceozone}\\
  \includegraphics[width=8.3cm]{pictures/final-total_ozone_drydepvelo}\\
  \includegraphics[width=8.3cm]{pictures/final-total_ozone_drydep}
  \caption{Comparison of the manifold of OsloCTM3 integrations with respect to (a) Surface ozone concentrations, (b) Annual average ozone dry deposition velocity, (c) Total annual ozone dry deposition. The different colors indicate sets of simulation with similar baselines. The multi-model mean from the evaluation of TF~HTAP models by \citet{ACP:Hardacre2015} is shown as a reference (where available).}
  \label{fig:mmm_drydep}
\end{figure}
%
In Appendix Fig.~\ref{fig:mmm_drydep_season}a), the average zonal ozone dry deposition is shown separated by month. Where available, we have added the multi-model-mean given by \citet{ACP:Hardacre2015} as reference. There are two major phases apparent: NH and SH greening season. Spring and summer in the NH is reflected in a pronounced peak of $v_\text{DD}^\chem{O_3}$ in the northern mid latitudes, while it is absent in winter (SH summer). Spring and summer in the SH are marked by a southward shift of the tropical peak dry deposition velocity and a slight increase of $v_\text{DD}^\chem{O_3}$ in the region $(20-40)\,\unit{^\circ S}$. In the Wesely scheme, NH mid latitude peak velocities appear in June compared to July in the EMEP scheme, indicating that the seasonal cycles differ. The corresponding total monthly ozone dry deposition is shown in Appendix Fig.~\ref{fig:mmm_drydep_season}b). In general, the seasonal patterns are quite similar in the Wesely scheme and the EMEP scheme, displaying a strong symmetry around $10\,\unit{^\circ N}$ in January/February and November/December, respectively. What differs most is the molding and intensity of the NH peak dry deposition. Both schemes reach the maximum in June/July but the peak is much more differentiated in March already in the Wesely scheme. Similarly, the SH tropical peak dry deposition is reached in August/September but sustained longer, into October, in the Wesely scheme. Since we have not conducted any simulation with a meteorological year other than 2005, we cannot elaborate on whether this is a special feature of our chosen year or not. In comparison, the available two month of the multi-model-mean display a different pattern. Especially the NH/SH asymmetry is much more pronounced in February, than it is the EMEP scheme.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Average seasonal cycles}
\label{subsubsec:seasons}
%
To further disentangle the contributions of different regions to the global ozone budget, we will briefly look at different projections of seasonal cycles.

@@

In Fig.~\ref{fig:mmm_drydep_season_pft}, the seasonal cycles of dry deposition velocities are shown for the nine surface categories. The patterns and absolute numbers differ substantially between the Wesely scheme and the EMEP scheme and the multi-model-mean. The divergence of the average dry deposition velocities between the Wesely scheme and the EMEP scheme in desert regions ($\Delta \overline{v}_\text{desert}^\chem{O_3} = 0.18\,\unit{cm\,s^{-1}}$) as well as grassland ($\Delta \overline{v}_\text{grassland}^\chem{O_3} = 0.54\,\unit{cm\,s^{-1}}$) is quite remarkable as well as the difference of both to the multi-model-mean in tropical forest regions ($\Delta \overline{v}_\text{tropical forest}^\chem{O_3} = 0.49\,\unit{cm\,s^{-1}}$). In contrast to the Wesely scheme and the EMEP scheme, the multi-model-mean indicates a rather pronounced seasonal cycle in desert regions ($0.10\,\unit{cm\,s^{-1}} \leq v^\chem{O_3}_\text{desert}\leq 0.15\,\unit{cm\,s^{-1}}$). The estimated dry deposition velocities over desert regions are consistent with the previously mentioned baseline values of ozone dry deposition, which means that in the EMEP scheme they are about one order of magnitude lower than in the Wesely scheme. From measurements of ozone fluxes in the Sahara desert, \citet{AE:Gusten1995} deduced $v^\chem{O_3}_\text{desert, day} = 0.1\,\unit{cm\,s^{-1}}$, $v^\chem{O_3}_\text{desert, night} = 0.04\,\unit{cm\,s^{-1}}$, and $\overline{v}^\chem{O_3}_\text{desert} = 0.065\,\unit{cm\,s^{-1}}$. This implies, that ozone dry deposition over desert regions is highly overestimated in the Wesely scheme as well as in TF~HTAP models, while it may be underestimated in the EMEP scheme. Similarly, the dry deposition velocities over water differ. From measurements over the ocean a mean value of $\overline{v}^\chem{O_3}_\text{water} = 0.019\,\unit{cm\,s^{-1}}$ has been deduced \citep{JGR:Helmig2012}. In a model study of different mechanisms of dry deposition to ocean waters by means of prescribed $\overline{v}_\text{water}^\chem{O_3}$ and one- and two-layer gas exchange modeling, \citet{ACP:Luhar2017} found $\overline{v}_\text{water}^\chem{O_3}$ ranging between $0.018\,\unit{cm\,s^{-1}}$ (two-layer scheme) and $0.039\,\unit{cm\,s^{-1}}$ (prescribed). With $\overline{v}^\chem{O_3}_\text{water} = (0.046\pm0.002)\,\unit{cm\,s^{-1}}$, the EMEP scheme (Section~\ref{subsubsec:Rb}) yields probably a too strong dry deposition to ocean which implies that ozone concentrations might even become larger and dry deposition even lower in the model if a two-layer scheme would be implemented. With respect to vegetation, we might be able to improve the model performance further by allowing more PFTs, especially in regions covered by tropical or coniferous forest. 

@@

Finally, we take a look at the different global as well as hemispheric sinks for ozone. Despite its vastness, the ocean amounts only to about $23\,\unit{\%}$ of the global ozone sink in the OsloCTM3 with operative EMEP scheme, while ice and snow account for far less than $1\,\unit{\%}$. The remainder is deposited to various other land surfaces of which deserts might be the most neglected in process modeling. The total annual dry deposition in the EMEP scheme is about $\frac{1}{3}$ below the multi-model-mean result by \citet{ACP:Hardacre2015}. But also the results of ~\citet{ACP:Luhar2017} yield a $(19-27)\,\unit{\%}$ lower ozone dry deposition than the models participating in the model intercomparison, with deposition to ocean ranging between $(12-21)\,\unit{\%}$.
%
\begin{table*}[t]
\caption{Total ozone dry deposition for the respective year in \unit{Tg\,a^{-1}} separated into ocean, ice and, land contributions.}
\begin{tabular}{lccccccccccccr}
\tophline
\multirow{3}{*}{Simulation} & \multicolumn{3}{c}{Ocean} & \multicolumn{3}{c}{Ice} & \multicolumn{3}{c}{Land} & \multicolumn{3}{c}{Total} & $\Delta^\dagger$\\
& NH & SH & Global & NH & SH & Global & NH & SH & Global & NH & SH & Global\\
& \multicolumn{3}{c}{(\unit{Tg\,a^{-1}})} & \multicolumn{3}{c}{(\unit{Tg\,a^{-1}})} & \multicolumn{3}{c}{(\unit{Tg\,a^{-1}})} & \multicolumn{3}{c}{(\unit{Tg\,a^{-1}})} & (\unit{\%})\\
\middlehline
Wesely\_type & 88.7 & 110.3 & 199.0 & 0.7 & 5.0 & 5.7 & 346.1 & 153.5 & 499.6 & 612.4 & 347.9 & 960.2 & 46.8\\
EMEP\_full & 67.1 & 84.7 & 151.8 & 0.6 & 4.4 & 5.0 & 236.0 & 112.3 & 348.2 & 408.6 & 245.6 & 654.2 & 0.0\\
EMEP\_offLight & 66.6 & 84.1 & 150.7 & 0.6 & 4.4 & 5.0 & 267.8 & 129.8 & 397.6 & 451.3 & 267.2 & 718.6 & 9.8\\
EMEP\_offPhen & 67.1 & 84.6 & 151.7 & 0.6 & 4.4 & 5.0 & 239.9 & 116.0 & 355.9 & 413.9 & 249.9 & 663.7 & 1.5\\
EMEP\_SWVL4 & 68.1 & 86.1 & 154.2 & 0.6 & 4.5 & 5.1 & 241.8 & 115.0 & 356.8 & 417.2 & 250.7 & 667.9 & 2.1\\
EMEP\_ppgs & 67.1 & 84.6 & 151.7 & 0.6 & 4.4 & 5.0 & 238.7 & 112.6 & 351.3 & 411.4 & 246.0 & 657.4 & 0.5\\
EMEP\_ppgssh & 67.0 & 84.5 & 151.6 & 0.6 & 4.4 & 5.0 & 243.9 & 119.4 & 363.3 & 419.4 & 253.6 & 673.0 & 2.9\\
EMEP\_ppgssh\_ice & 66.5 & 84.8 & 151.3 & 0.2 & 1.3 & 1.4 & 240.3 & 120.4 & 360.7 & 413.0 & 251.1 & 664.1 & 1.5\\
EMEP\_ppgs\_2005 & 66.8 & 83.5 & 150.4 & 0.6 & 4.4 & 5.0 & 237.8 & 110.0 & 347.8 & 409.2 & 240.7 & 649.9 & -0.7\\
\bottomhline
\end{tabular}
\belowtable{$^\dagger$: Difference of global annual total relative to \emph{EMEP\_full}.} % Table Footnotes
\end{table*}
%

@@

In Fig.~\ref{fig:mmm_drydep_stations}a), seasonal cycles of average ozone dry deposition fluxes for six observation sites are shown. We have computed a model average for all sensitivity studies at the closest grid point and show the $1\,\sigma$ error band. The shaded area around the multi-model-mean indicates the broad range of model results but is not an actual error band since such is not given in \citet{ACP:Hardacre2015}. At about $4$ of $6$ sites, the EMEP scheme performs better than the Wesely scheme and better than or the same as the multi-model-mean. We use a $\chi^2$-test
\begin{equation}
  \chi^2 = \sum_{i=1}^{12}\frac{\left(\overline{\chem{O_3}}_{\text{DD,\,}i}^\text{sim}-\overline{\chem{O_3}}_{\text{DD,\,}i}^\text{obs}\right)^2}{\sigma_i^2},
\end{equation}
with an estimated standard deviation of observation $\sigma_i=1\,\unit{mmol\,m^{-2}\,s^{-1}}$ and divide it by the number degrees of freedom (NDF) to assess this subjective analysis in a more objective way. The closer to $1$ this test scores, the better does the simulation represent the observation. A score between $0$ and $1$ indicates that the estimated $\sigma$ is too small. The results of the $\chi^2$-test are shown together with the divergences in Fig.~\ref{fig:mmm_drydep_stations}. The $\chi^2$-test reveals also that in $4$ of $6$ cases the EMEP scheme improves the performance of the OsloCTM3 with respect to ozone dry deposition fluxes, although a satisfying result is only achieved for two sites (Castel Porziano, Blodgett Forest).
With only one full year of simulation, the model uncertainty regarding the seasonal cycle at observational sites cannot be quantitatively assessed. Further more, the observational averages uses at most nine years worth of data which questions the statistical reliability of these data. Among other aspects that might affect the model performance the horizontal as well vertical resolution play an important role. In the rather coarse resolution, good matches between observation and model are only to be expected if the station is at a location which is representative for an area similar to the model resolution.

%
\begin{figure}[t]
  \includegraphics[width=8.3cm]{pictures/final-monthly_ozone_drydep_stations_True}\\
  \includegraphics[width=8.3cm]{pictures/final-monthly_drydep_divergence_True}
  \caption{Ozone dry deposition fluxes at different observation sites. (a) Comparison between Wesely scheme and average result from all sensitivity studies with observational averages taken from \citet{ACP:Hardacre2015}. The model uncertainty of the OsloCTM3 is given as $1\,\sigma$ error band. The shaded area around the multi-model-mean indicates the broad range of different model results but is not an actual error band. (b) Model divergence from observation and $\chi^2$-test results.}
  \label{fig:mmm_drydep_stations}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\conclusions[Discussion and conclusions]  %% \conclusions[modified heading if necessary]
\label{sec:conc}
We have presented an update of the dry deposition scheme in the OsloCTM3 from prescribed dry deposition velocities \citep{AE:Wesely1989,JGR:Hough1991} to a more process oriented parameterization taking the state of the atmosphere and vegetation into account \citep{ACP:Simpson2012}. The dry deposition to vegetation is now sensitive to parameters such as photosynthetic active radiation, water vapor pressure deficit, soil water content, and plant phenology. The update decreases the annual amount of ozone dry deposition up to $100\,\unit{\%}$ especially over large desert regions and the oceans. Compared to results from a multi-model evaluation \citep{ACP:Hardacre2015}, the total annual ozone dry deposition in the OsloCTM3 is about $33\,\unit{\%}$ below the average. However, there seems to be a tendency, that the newer TF~HTAP models show a lower total annual dry deposition of ozone than older models, indicating that newer developments in the models lead to decreasing ozone dry deposition and increasing tropospheric ozone burden \citep[e.g.,][]{ACP:Luhar2017,AE:Hu2017}.
